A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-19 tsibble.Rmd
done with sec 5 revision

The tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data. The supporting operations can be chained for sequencing analysis, articulating a data pipeline. As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline. The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package.

@@

The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then". A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.

@@

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal context, particularly useful when handling a large collection of data. The contextual summary provided by the print function, shown below on the data from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of series. These details aid users in understanding their data better.

@@

Rolling window uses a programming paradigm---functional programming, which is different from those table verbs listed in Table \@ref(tab:verb-table). Table verbs expect and return a tsibble, and does what the function name suggests. On the contrary, these rolling window functions could accept arbitrary input types and would return arbitrary sorts of output, depending on which method is put into the rolling window. For example, computing moving averages requires numerics and a function like `mean()`, and produces averaged numerics. However, rolling window regression takes a data frame and a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.

The **purrr** package [@R-purrr] provides a good example of functional programming in R. It provides a complete and consistent set of tools to iterate each element of a vector with a function. Rolling window does not just iterate but rolls over a sequence of elements, namely `slide()`, `tile()` and `stretch()`. `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs. For type stability, the functions always return lists. Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding type, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively. Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr]. This family of functions empowers users to incorporate window-related operations in their workflows.

@@

Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and manageable chunks, and provides developers with simpler maintenance.

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly deal with time. The atomic functions (such as `yearmonth()` and `yearquarter()`) embedded in the `index_by()` table verb collapses a tsibble to a less granular interval. The substitution of another time function in the `index_by()` results in the aggregation of different time resolution. Since these time functions are not exclusive to a tsibble, they can be used in a variety of applications in conjunction with other packages. The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions. It is also generally easier to trace back the errors users encounter from separating 1d and 2d functions, and increase the code readability.

@@

Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution implemented in **nanotime** [@R-nanotime], and pure numbers in simulations. Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday. The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters. These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year. A new index would be immediately recognized by the software upon defining `index_valid()`, as long as it can be ordered from past to future. The interval regarding semesters is further outlined through `interval_pull()`. As a result, the rest of the software methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underlying basis of temporal data, and there is a demand for sub-classing a tsibble. For example, a fable is actually an extension of a tsibble, mentioned in Section \@ref(sec:ts-models). A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. This new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, like prediction distributions to a fable. Tsibble verbs are also S3 generics. Developers will be able to implement these verbs for the new class, if necessary.

@@

The **tsibble** packages leverages the **tidyverse** grammars and pipelines through tidy evaluation [@tidy-eval] via the **rlang** package [@R-rlang]. In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**. This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive.

>>>>

To sum it up, the tsibble abstraction provides a formal organization of forwarding tidy data to model-oriented temporal data. The supporting operations can be chained for sequencing analysis, articulating a data pipeline. As @eopl stated, "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." A mini snippet below glues transformation and forecasting together, to envision the fluent pipeline.

@@

The `pedestrian` dataset [cite] contains hourly tallies of pedestrians at four counting sensors in 2015 and 2016 in inner Melbourne, available in the **tsibble** package. The pipe operator `%>%` introduced in the **magrittr** package [@R-magrittr] chains the verbs, read as "then". A sequence of functions are composed in a way that can be naturally read from left to right, which improves the code readability.

@@

The **tsibble** package development follows closely to the **tidyverse** design principles [cite].

@@

Beyond the tools, the print display provides a quick and comprehensive glimpse of data in temporal contexts, particularly useful when handling a large collection of data. The contextual information provided by the `print()` function, shown below from Table \@ref(tab:tb-sub), contains (1) data dimension with its shorthand time interval, alongside time zone if date-times, (2) variables that constitute the "key" with the number of units. These summaries aid users in understanding their data better.

@@

Rolling window adapts to functional programming, which the **purrr** package [@R-purrr] sets a good example. These functions accept and return arbitrary inputs and outputs, with arbitrary methods. For example, moving averages anticipate numerics and produce averaged numerics via `mean()`. However, rolling window regression feeds a data frame into a linear regression method like `lm()`, and generates a complex object that contains coefficients, fitted values, and etc.

Rolling window not just iterates but rolls over a sequence of elements of a fixed window. A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants. `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs. For type stability, the functions always return lists. Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively. Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].

@@

Modular programming is adopted in the design of the **tsibble** package. Modularity benefits users by providing small focused and cleaner chunks, and provides developers with simpler maintenance.

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly operate on time. The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval. Since they are independent from a tsibble, they can be used in a broader range of data applications not constrained to tsibble. On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.

@@

Time representation could be arbitrary, for example R's native `POSIXct` and `Date` for versatile date-times, nano time for nanosecond resolution in **nanotime** [@R-nanotime], and numerics in simulation. Ordered factors can also be a source of time, such as month names, January to December, and weekdays, Monday to Sunday. The **tsibble** package supports an extensive range of index types from numerics to nano time, but there might be custom indexes used for some occasions, for example school semesters. These academic terms vary from one institution to another, within an academic year which is defined differently from a calendar year. A new index would be immediately recognized upon defining `index_valid()`, as long as it can be ordered from past to future. The interval regarding semesters is further outlined through `interval_pull()`. As a result, all tsibble methods such as `has_gaps()` and `fill_gaps()` will have instant support for data that contains this new index.

The class of tsibble is an underpinning for temporal data, and sub-classing a tsibble will be a demand. A low-level constructor `new_tsibble()` provides a vehicle to easily create a new subclass. This new object itself is a tsibble. It perhaps needs more metadata than those of a tsibble, that gives rise to a new data extension, for example prediction distributions to a forecasting tsibble.

@@

The **tsibble** packages leverages the **tidyverse** grammars and pipelines through tidy evaluation [@tidy-eval] via the **rlang** package [@R-rlang]. In particular, the table verbs extensively use tidy evaluation to evaluate computation in the context of tsibble data and spotlights the "tidy" interface that is compatible with the **tidyverse**. This not only saves a few keystrokes without explicitly repeating references to the data source, but the resulting code is typically cleaner and more expressive, when doing interactive data analysis.
